Clouds and Sensitivities Across a Hierarchy of GFDL CMIP6 Models

### 2020-11-03 Silvers_WalkerCell_response2reviewers.tex

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

%% ------------------------------------------------------------------------ %%

\section{Note from LGS and TR}

\textcolor{blue}{All responses and comments by the authors are in blue.  Below please find my responses to your particular considerations.  When referring to line numbers that were used in your revues I have done my best to use the same line numbers (those from the original submission) unless otherwise
noted.  Reviewer \#1 made additional comments in a pdf that was included with the review.  I have imported those comments to this document.  Reviewer \#3 made 15 minor comments in the annotated pdf provided by \textit{JAMES}.  I have responded with my own annotations in the annotated pdf. }

@@

This paper presents a number of interesting results on the interactions between longwave cloud
radiative effects and the Walker circulation using GCM-like and CRM-like simulations. This
approach provides an important link between more idealized, smaller-scale simulations of the
tropical atmosphere where convection is resolved (i.e. RCE with a CRM) with less-idealized,
larger-scale simulations of the tropical atmosphere where convection is not resolved (i.e.
prescribed SST gradients with a GCM). As such, this work is critical for improving our
understanding of the observed tropical atmosphere.

The central finding is that the CRM-like simulations produce substantially less low clouds than
the GCM-like simulations, and this leads to important differences in the atmospheric response to
longwave cloud radiative effects (LWCRE). In particular, LWCRE act to increase precipitation
in the GCM-like experiments, and act to decrease the precipitation in the CRM-like experiments.
The results have important implications for our understanding of the interactions of clouds and
the tropical circulation in nature as well as in more sophisticated global climate simulations.

In my view the main weakness of the paper is lack of clarity in a few areas. The paper could
benefit from emphasizing the main points, as well as clarification of some of the science ideas.
The figures generally illustrate the main points nicely, but I have included a few notes for
improvements, as well as suggestions for additional plots (particularly a plot of the longwave
heating rate as a function of distance). See attached document for comments*, which have been
highlighted throughout the paper. The suggestions are generally minor, but I think the clarity of
the paper could be improved substantially with these changes.

\textcolor{blue}{Thank you for your review and the encouraging synopsis, it has helped me to improve the paper.  I think that 
the changes made as a result of the comments from you and the other reviewers has substantially improved the clarity of the 
results and descriptions.  Your comments have helped to better emphasize the main points and to improve the representation
of the data with the figures.}

@@

   \textcolor{blue}{We have taken your advice, mostly by moving the content from the former second to last sentence in the abstract to earlier in the abstract and rewriting lines 27-28 of 
   the originally submitted manuscript.}

\item Ln 67-68:  The Walker circulation itself is not really a framework, and comparing it to observations does not really follow since it occurs in reality. I think you mean to say that idealized or ‘mock-Walker’ circulations can be used as a framework to understand the interactions between clouds and circulation, both in observations and in comprehensive climate model simulations? 
  \textcolor{blue}{I agree the Walker circulation itself is not a framework.  The sentence in question now reads, ``The Walker Circulation provides an excellent example with which to compare the overturning circulation and interactions between clouds and radiation using a variety of model configurations.''}

\item Ln 71: Change to lowercase: 'sea surface temperature'  
  \textcolor{blue}{Done}

  \item Ln 119-120: Best to avoid colloquial phrases. I gather that you are saying that it is difficult to compare, for example, a CRM with a large domain, to a GCM, because there the model physics differ in many ways, not just how convection is treated. So differences between those models are not necessarily due to the convection alone. Your work helps clarify these issues by using the same model.
  \textcolor{blue}{The text now reads: Current climate models continue to be developed with an increasingly fine resolution and 
the domain sizes used with CRMs continues to grow.  The parameter space of GCMs and CRMs has begun to overlap but the representation of 
physics continues to differ by much more than grid-spacing and the treatment of convection.  As a result there is a need to systematically compare the 
clouds and their influence on the climate produced by each type of model.
By simulating a mock-Walker Circulation in the context of both a GCM and a CRM that have been derived from the same model we 
illustrate how inextricable the interactions between clouds and radiation are to the coupling of moisture with the large-scale circulation.}

  \item Ln 123-124: Can you be more specific here?
  \textcolor{blue}{Please see my response to your previous comment, on Ln 119-120, I trust this has clarified the statement for readers.}

@@

  \item Ln 216-218: It seems to me that it would be better to refer to any experiment in the range of P100-P25 as 'GCM-like' and any experiment in the range of E25-E1 as 'CRM-like'. Then, the E25 and P25 experiments serve as the link between GCM-like and CRM-like experiments.
  \textcolor{blue}{There was a mistake in this sentence, it has been corrected to include, ``...`GCM-like' while the E2 and E1
  experiments are referred to as CRM-like''. Additionally, you make a good point here, that the distinction between GCMs and CRMs could be made based on whether a 
  parameterization is used for the convection.  However, in my opinion, a model with 25km grid-spacing should not be called cloud-resolving.  I also considered the possibility of referring to the two classes of model-types as 'explicit models' and 'parameterized models' as a way to more clearly distinguish them.  But I chose against that naming convention because I think it then blurs the natural similarities that my P100, P25, E25 simulations have with GCMs and the natural similarities my E1, E2 simulations have with CRMs.  Using parameterized and explicit in the naming also overly emphasizes the convective parameterization despite all simulations involving many identical parameterizations.}

  \item Table 2:  This table is helpful. I wonder if adding two columns for high cloud and low cloud fraction could help further summarize some of your key results?
  \textcolor{blue}{Good suggestion.  I have made several additions to the paper that I think further clarify the results as related to the cloud and moisture 
  distributions with height.  The new Figure 4 shows profiles of cloud fraction and cloud condensate for the P100, P25, and E25 experiments.  I have also 
  added a panel to Figure 13 which  shows the profiles of specific humidity.}

  \item Figure 1: Nice figure. It is interesting that the middle and right panels are so similar to each other. Does this suggest that the cloud longwave effects are weak in the E25 simulation? And potentially much too strong in the P25 simulation? It might be useful to show a plot of the longwave heating rate for each simulation...  
  \textcolor{blue}{I agree, the similarity of panels b and c is interesting.  However, I hesitate to conclude that LW CRE is very weak in E25 LWCRE-on and 
  too strong in the P25 simulation.  Mostly because I don't know what constitutes a LW CRE that is `too strong'.  I interpret the difference between panels
  a and b to indicate how important the LWCRE is to the atmospheric state, and the symmetry of panel c to indicate the more regular behavior of explicit convection.
  While panel b is much more symmetric than panel a, there are still notable asymmetries present which supports the idea that part of the asymmetry comes from the convective parameterization and part of it comes from the LWCRE.  Figure 5 of the revised manuscript, with five year Hovmoller diagrams for the cases in Figure 1 will be interesting for you to see.  I have almost added plots of the 2D longwave heating many times to this paper.  However, they don't add much insight and the paper is already bordering on unwieldy, so I have refrained.}

@@

  \item Ln 282-284:  Doesn't Fig. 1 show that P25 control yields a drier atmosphere? You also state this in lines 278-280. I guess it yields a drier subsiding region, but also a much more humid convective region. 
  \textcolor{blue}{Yes, the end of your statement is correct, the subsiding region is drier, but the convective region is more humid than the other experiments.  This is stated in the lines you mention, as well as directly in lines 276-278.  It is the case that the mean values are a bit different than one might expect from looking at Fig. 1.  This also highlights the difference between the means of water vapor compared to water condensate.}

  \item Ln 310: I only see prominent asymmetry in the P25 simulations. The P100 simulation is only slightly asymmetric, and the P100L simulation not asymmetric at all.
  \textcolor{blue}{The language describing this (in multiple places) has been changed to reflect off center features, rather than just asymmetric.  Please see the discussion in 
  the new subsection (3.2) which discusses the comparison between P100, P25 and E25.   While P100 is only slightly asymmetric, it is 
  consistently off center.  The asymmetry of P25L is greater than may be initially apparent because of the difference in domain size.  I am happy to change 
  the language here as I don't think the 'asymmetry' is important scientifically.}

  \item Ln 321-323:   If I'm understanding this correctly, it seems that Fig. 1 show that it is mostly the interaction between LWCRE and 
  circulation that drives this asymmetry. Because Fig 1. middle panel is very symmetrical (with LWCRE disabled).
Also, I wonder to what extent the domain shape influences 
  this result (the asymmetry is not present in P100L and less prominent in P25L)
  \textcolor{blue}{The new subsection, 3.2, and especially the new Figure 5, will clarify this further.  It should be pointed out that while the middle panel of Fig1 is symmetrical relative to the left panel, 
  the far right panel is even more symmetrical (at least in terms of the RH and streamfunction).}

  \item Ln 338-343: Yes, but maybe clarify that you must compare similar domain sizes in Table 2. The variation in SF is also influenced by domain size. 
  \textcolor{blue}{I agree in principle that the cleanest comparisons will be with similar domain sizes, but comparing similar domain sizes does not help to clarify the numbers in Table 2.  For example, both the largest and smallest SF are found on the large domain.  I also think there is still some value 
  in the comparison even across different domain sizes.}

@@

  \item Figure 6: I think there are too many lines on this plot. Maybe two separate plots would be best: one for P25 and one for P100.
  \textcolor{blue}{Done.  Splitting this into two panels was a good suggestion.}

  \item Ln 461: 'more' instead of 'better'
  \textcolor{blue}{Done}

  \item Figure 13: These plots are useful, however I still think it might be helpful to show a profile of the longwave heating rate vs. horizontal distance (like Fig. 10)
%
Also, the triangle markers for E25 may be obscuring some details. 
I wonder if just using a slightly different color for E25 and P25 would be better.
\textcolor{blue}{Figure 13 has been redone as a 2x2 panel figure using a different color for E25 instead of the triangles, larger axis labels, and clearer panel 
titles.  I have opted to not include figures of the longwave heating vs horizontal distance.  I have not found them to provide unique insight and are best 
understood in combination with the shortwave heating, convective heating, and heating due to large-scale physics plots.  Including this all would significantly 
bulk up the paper with minimal gain.}

@@

  \item Ln 545-546: This is particularly true because the low clouds are found in relatively dry regions. Without low clouds, the dry atmosphere would emit less downwelling longwave radiation and roughly the same outgoing longwave radiation. In other words, the low clouds effectively increase the downwelling radiation in dry regions, which leads to more atmospheric cooling.   
  \textcolor{blue}{This is a good way of stating the underlying reasons.  I have made some modifications so the sentence now 
  reads, ``Low-level clouds, especially in dry regions, serve as a significant source of  radiative cooling for the atmosphere (Figure 
  13) because they increase the downwelling longwave radiation.''}

  \item Ln 556-557: A bit unclear. What decrease of low-level clouds are you referring to here? Are you referring to the fact that the CRM-like simulations have fewer low clouds?
  \textcolor{blue}{I have added, `when the LWCRE is off' to this sentence to make my meaning clear.  I am not referring to the fact that CRM-like simulations have fewer clouds, but to the fact that there are far less low-level clouds in the LWCRE-off simulations 
  especially for the GCM-like simulations. }

  \item L584-608: I think it is good to have a summary here, but I'm getting lost in all of the details. It would be helpful for the reader if you could summarize only the most important points (maybe 5-6 bullets max). For example, I think you could probably eliminate the last 2 points in the first set (i.e. about spatial gradients of humidity, etc and about the condensate changes).

You also do not summarize any of the domain size results here, however they are mentioned in your conclusions. So I'm left a bit confused about the overall importance of the domain size results... to remedy that I suggest 1) mentioning the key domain-size sensitivies here and/or 2)  deleting the discussion of the domain-size sensitivities in the conclusions. Just trying to help you better convey the main points to the reader.
  \textcolor{blue}{The last bullet point (Lines 592-594) of the first list has been removed.  I have also reorganized and combined the material that was 
  formerly in sections 6 and 7 into one section with the hope of more clearly distinguishing between the summary of our results and the discussion points.  Hopefully it is not as easy to get lost in the details now.}

  \item Ln 690-692: Can you speculate about why this might be? Could the smaller low-cloud amounts in the CRM-like experiment be due to the stronger overturning circulations, and hence stronger subsidence drying? 
  \textcolor{blue}{I have included the following in the freshly organized concluding section: 
  Our results show a stronger of the overturning circulation (Figures 1,2, and 11)
 in the CRM-like experiments compared to the GCM-like experiments.  
 Accompanying this is a difference in the vertical mixing of water condensate, specific
 humidity and temperature (Figures 4, 8, and 13).  We conclude that the CRM-like experiments (E1 and E2) transport more specific humidity 
 and liquid water from the lowest 200 hPa of the atmosphere into the troposphere above than do the GCM-like experiments 
 (P100 and P25).  While the E25 experiment does not use convective parameterizations, in terms of the vertical distribution of water
 species and clouds it is more similar to the GCM-like experiments than to the CRM-like experiments.}

@@

  \item Ln 701: 5\%
  \textcolor{blue}{Done}

  \item Ln 704: context 'of' 
  \textcolor{blue}{Done}

  \item Ln 704-706: I know this is not the focus of your study, but I wonder which simulation has the more realistic cloud profile: the CRM or GCM-like simulation? The low-cloud fraction seems quite small in the CRM-like experiments...
  \textcolor{blue}{I have added a new figure that looks at observations from the tropical Pacific in order to give a sense of how these model results compare to observations.  
  Figure 15 shows cloud fraction from the MISR instrument on board NASA's Terra satellite.  I focused on a region that is usually identified with the Walker Circulation and plotted profiles for the domain mean, a region with upwelling, and a region with downwelling.  If comparing to figure 14 in our manuscript in which we look at the cloud fraction in the subsidence 
  region it is seen that both the upper-level and low-level cloud fraction observed by MISR in the downwelling region is closer to that of the GCM-like simulations.  This is 
  somewhat surprising but it should be kept in mind that this type of comparison is very qualitative and subject to numerous caveats.  For example, impacts from the Hadley cell 
  have not been filtered out of this observational data.  I am hopefully that a quick look at data like this could help to motivate future studies in which a more detailed comparison 
  between models of the Walker Circulation and observations.  The following text has been added to the manuscript:}

  \textcolor{blue}{In order to make a qualitative connection to observed clouds and hopefully motivate a more detailed comparative 
study we look 
at the clouds observed with the Multiangle Imaging Spectro-Radiometer (MISR) instrument on board NASA's Terra satellite (see 
Marchand et al., 2010 for details).  We use the domain (10N-35S) which was identified by Schwendike et al., 2014 as defining the 
regional Walker circulation in the Pacific.  The left panel of Figure 16 shows the cloud fraction averaged from March 2000-
November 2019 of the full domain of the Pacific regional Walker circulation.  This region includes much of the so-called warm pool 
beneath the ascending branch of the Walker Circulation as well as the eastern Pacific subsidence regions that encompass the 
stratocumulus cloud decks that are common off the west coast of South America.   To calculate profiles of the cloud fraction as a 
function of height we sum the MISR data over all optical depths  and plot the domain mean (dashed line), a region of subsidence 
($10^\circ$ N-$35^\circ$S; $120^\circ$W-$60^\circ$W, solid line), and a region that is characterized by deep 
convection ($10^\circ$N-$35^\circ$S; $120^\circ$E-$180^\circ$, dash-dot line).  Overall, both the upper- and low-level cloud fraction 
from MISR are closer to the cloud profiles simulated by the GCM-like models.  However, it should be kept in mind that there are 
several substantial differences between our mock Walker circulation and the observed Walker circulation.}

@@

This manuscript examines a simulated Walker circulation-like setup, making use of a flexible model that can be configured into GCM-like and CRM-like states. By focusing primarily on the equilibrium state attained in each simulation, the coupling of clouds, radiation, and circulation is studied. These each change substantially as domain size, resolution, use of a convective parameterization, and interactions between clouds and longwave radiation are altered separately. Results also have potential implications toward studying the complex cloud radiative feedbacks that affect our future climate projections.

The manuscript offers a thorough analysis of this intricate coupling, and will undoubtedly be a significant contribution to the literature. As you will see below, the bulk of my comments relate to the organization of the manuscript, which I believe should be revisited and can be improved to present these meaningful results in a more cohesive and concise manner. In addition, I believe more attention should be given to the comparison between the two 25 km-resolution simulations: those with parameterized convection and explicit convection, respectively. While these suggested organizational edits may be more of a significant undertaking, I ultimately believe this set of revisions to be more minor, and look forward to reading the finished product!

Below, I have organized my comments into a set of general, overarching points first, followed by a handful of individual, line-by-line comments.

\textcolor{blue}{Thank you for your review and the encouraging synopsis, it has helped me to improve the paper.  I think that 
the changes made as a result of the comments from you and the other reviewers has substantially improved the clarity of the 
results and descriptions.  I have incorporated most of you recommendations and trust that you will find the manuscript substantially 
improved.}

@@

\subsection{GENERAL COMMENTS}
*In terms of organization, I believe some reworking and consolidation of results are useful and fairly straightforward. Some examples include:

### 2020-11-03 Silvers_cfmip.tex

This paper has a double motivation.  First, to illustrate the wide range of climate sensitivities to external perturbations 
that can be achieved with a single base model.  For this purpose we utilize several models from the new suite of GFDL 
models that participated in CMIP6.  This constitutes a hierarchy of configurations that ranges from an atmosphere only 
aquaplanet to a fully coupled 
climate model.  Previously the range of climate sensitivities and feedbacks among the ensemble of climate models 
the world has to offer have been documented largely by examining the multi-model mean results with less attention to 
the variability of sensitivity within a single modeling framework (IPCC; \citeA{Andrews_etal_2012, Ringer_etal_2014}).  By 
focusing on the models from a single modeling center we seek fresh insight that could be buried within a large ensemble.

@@

A second motivation for this paper is to document the simulation of clouds in the newly developed GFDL models that 
have participated in CMIP6.  This will largely be done in the context of the amip experiment and comparisons to observations
that overlap with parts of that time period (1979-2014).  The experiments and diagnostics used are taken from the latest 
round of CFMIP which has provided a useful framework with which to evaluate cloud feedbacks.

@@

Based on constant temperature perturbation experiments CM4 was expected to be significantly less sensitive 
than CM3 \cite{Zhao_2014}.
Recent research has made it clear that the climate sensitivity estimated from constant temperature perturbation 
experiments is not a good indicator of the climate sensitivity that is estimated from regression techniques of 
abrupt 4xCO2 warming experiments, and that both 
of these measures of sensitivity underestimate the equilibrium sensitivity that is attained when a coupled climate 
model reaches equilibrium after multiple millennia \cite{Paynter_etal_2018,Rugenstein_etal_2020}.     
At least part of the reason for this multiplicity of 
sensitivities obtained from different method of calculation is that variations of the climate sensitivity within 
individual models has been tied to the influence of particular surface temperature patterns 
\cite{Gregory_Andrews_2016, Zhou_etal_2016, Andrews_Webb_2018, Silvers_etal_2018}.   
The influence of a pattern in the surface temperature is communicated to the TOA radiative fluxes by the clouds 
(Gregory and Andrews, 2016; Rose et al.; Silvers et al., 2018).   Thus, clouds are important factors to both the uncertainty 
estimates of climate sensitivity, as well as to the variability in time of the climate sensitivity.

This work assesses the fidelity of the clouds in AM4.0/CM4.0 with observations and compares clouds from experiments 
with differing sensitivities. 
Section two compares the clouds from the \textit{amip} run of AM4.0 with satellite observations.  That section also 
analysis the radiative fluxes at the top-of-the-atmosphere (TOA) and compares them to the observations from 
CERES-EBAF.    Section three documents the climate sensitivities of GFDL models across multiple CMIP6 experiments 
and discusses the potential role played by clouds.   Conclusions are drawn in section four.

@@

We compare output from the AM4.0 GCM to observations from ISCCP, MISR, MODIS, CALIPSO, and CERES-EBAF.
More details are provided in the appendix.  The CMIP6 and CFMIP experiments used in this study are given in Table
\ref{tab:exps} and the observational data sets used are given in Table \ref{tab:obs}.  Additional details about all 
experiments used in this paper are given in \cite{Eyring_etal_2016} and \cite{Webb_etal_2017}.  The experiments are referred to with the
appropriate CMIP6 labels.

The comparisons with observations are made mostly with the \textit{amip} experiment.   The experiments used 
in this paper are all part of the GFDL contribution to CMIP6 and as a result, the \textit{amip} and \textit{historical}
experiments end in the year 2014.  We use the \textit{historical, abrupt-4xCO2}, and \textit{1pctCO2}
experiments from the CMIP6 DECK.  In addition we utilize several experiments from the CFMIP contributions 
to CMIP6.  These experiments include \textit{amip, amip-p4K, amip-m4K, amip-future4K, aqua-control} and \textit{aqua-p4K}.  
In order to utilize as much of the available data as possible much of the analysis of the \textit{amip} experiment uses the 
full span of years (1979-2014) despite the fact that the satellite observations do not extend as far back in time.  However, 
some of the analysis for the \textit{amip} and \textit{historical} experiments, in particular for the calculation of a climate feedback parameter over recent years, a shorter period (1996-2014) is used in order to eliminate complications due to radiative forcing in the early 1990's due to volcanic activity.  The relevant time period is stated in the captions of the figures and tables.

@@

Before discussing the climate sensitivities of AM4/CM4 and how they are influenced by cloud feedbacks we assess
the clouds and radiative fluxes which are simulated by AM4.0 over the recent period of 1979-2014.  This is done 
by comparing satellite based observations to the \textit{amip} experiment that is part of GFDLs 
contribution to CMIP6.  
During the development and documentation of AM4.0 \cite{Zhao_etal_2018a, Zhao_etal_2018b}, 
(Z18a and Z18b hereafter) 
and CM4 \cite{Held_etal_2019}, specific cloud properties such as the 
cloud droplet number, cloud fraction, and the partition of cloud water into liquid or ice were noted, but 
accurately modeling the radiative fluxes at the top of the atmosphere was determined to be a 
higher priority partly because of the large observational uncertainties of those cloud properties (Z18a,b).   
Based on the initial analysis for those papers, the cloud properties of AM4.0 were determined to be
at least as consistent with observations as previous generations of GFDL models.  Z18a also noted 
that AM4.0 underestimates the liquid droplet number over oceans and downwind of the aerosol emission 
sources.   
Comparisons of AM4.0 with observations from ISCCP and MODIS showed an underestimate of 
the optically thin low clouds and the mid-level clouds (Z18a, Figure 10).  
However, the observational uncertainty for some of these properties prohibited firm conclusions.  This paper
uses additional observations and a more thorough analysis of the clouds in AM4/CM4 to better determine 
the quality of the cloud simulations.        
In this section we first compare the simulated and observed cloud fraction.  We then compare  
the simulated TOA radiative fluxes from AM4.0 to those observed by CERES.    
Throughout the paper we utilize what has come to be known as the Cloud Radiative Effect (CRE), calculated 
as the difference between the full-sky and clear-sky radiative fluxes.  Earlier literature \cite<e.g.,>[]{Cess_etal_1989, Soden_etal_2004} often referred to the CRE as the Cloud 
Radiative Forcing (CRF).

The CRE is a simple and effective tool for evaluating the influence of 
clouds on the climate.  It is well known (Soden et al., 2004) that a change in the CRE is not a direct measure of the cloud 
feedback, and that non-cloud related values of CRE can occur.  However, there are many examples of 
the productive use of the CRE as a way to diagnose cloud effects (e.g. Cess et al., 1989; Webb et al., 2001; 
\citeA{Ringer_etal_2014}; Andrews et al., 2015), it is simple to calculate, and
it allows for direct comparisons with previous studies.  The pure cloud feedback usually corresponds 
to a positive shift of the change in CRE values by roughly 0.3 $\rm{W m^{-2}K^{-1}}$ (Soden et al., 2004).  The total cloud 
feedback of a model is therefore more positive than will be implied by the CRE.

@@

\subsection{COSP: change title}
In order to compare the observed and modeled clouds as consistently as possible across multiple platforms 
GFDL has implemented a collection of satellite simulators which are part of the 
CFMIP (Cloud Feedback Model Intercomparison Project; see \cite{Webb_etal_2017}).  Satellite simulator
are software tools that are used in coordination with GCMs to allow for model variables 
to be translated to variables that are more directly comparable with what satellites observe.  
These simulators are often referred to as the CFMIP Observation Simulator Package (COSP); for
details see Bodas-Salcedo et al. 2011.  This paper highlights simulators of the algorithms for 
four satellite datasets: ISCCP (the International Satellite Cloud Climatology Project), 
MODIS (the Moderate Resolution Imaging Spectroradiometer), 
MISR (the Multiangle Imaging Spectro-Radiometer), and 
CALIPSO (the Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation; 
Chepfer et al., 2010; Cesana and Chepfer, 2013).  By using multiple complimentary satellite 
datasets and simulators we can benefit from the different strengths of each satellite while largely 
minimizing the known problems and discrepancies of the observational data sets.

@@

The MISR satellite provides good observations of low-level cloud heights and amounts.  
For this reason and for comparison with both Kay et al., 2012, and Zhang et al., 2019 we compare the cloud fraction 
between observations from MISR and the AM4.0 simulator output for clouds below 3km in Figure \ref{fig:cf_misr_lowcl}.  
The optical depth is used to distinguish between the thick and thin clouds.  
Consistent with comparisons with ISCCP and CALIPSO we see that AM4.0 has fewer clouds than observed.
For thick low-level clouds the negative cloud fraction bias is about 14\%, and 12\% for thin clouds.   
For thick clouds AM4.0 does reasonably well producing the observed spatial distribution of clouds, except in the tropical Indian Ocean. 
The pattern of thin clouds is not reproduced as well, particularly in the southern hemisphere and the tropics.

@@

>>>>

This paper has a double motivation.  First, to illustrate the wide range of climate sensitivities to external perturbations 
that can be achieved with a single base model.  For this purpose we utilize several models from the new suite of GFDL 
models that participated in CMIP6.  This constitutes a hierarchy of configurations that ranges from an atmosphere only 
aquaplanet to a fully coupled global
climate model.  Evaluating the range of climate sensitivities and feedbacks among climate models 
has previously been documented most commonly by examining the multi-model mean results (IPCC; \citeA{Andrews_etal_2012, Ringer_etal_2014}) with less attention to 
the variability of sensitivity within a single modeling framework.  
The exception to this multi-model ensemble approach is to examine the impact of a large number of parameter perturbations
within a single model (referred to as a Perturbed Physics Ensemble, see Murphy et al., 2004; Stainforth et al., 2005; Klocke et al. 2011).  
In this paper we focusing on a single base model across a range of different warming scenarios and model configurations.

@@

@@

Based on constant temperature perturbation experiments CM4 was expected to be significantly less sensitive 
than CM3 \cite{Zhao_2014}.
Recent research has made it clear that the climate sensitivity estimated from constant temperature perturbation 
experiments is not a good indicator of the climate sensitivity that is estimated from regression techniques of 
abrupt 4xCO2 warming experiments, and that both 
of these measures of sensitivity underestimate the equilibrium sensitivity that is attained when a coupled climate 
model reaches equilibrium after multiple millennia \cite{Paynter_etal_2018,Rugenstein_etal_2020}.     
At least part of the reason for this multiplicity of 
sensitivities obtained from different methods of calculation is that variations of the climate feedback within 
individual models has been tied to the influence of particular surface temperature patterns 
\cite{Gregory_Andrews_2016, Zhou_etal_2016, Silvers_etal_2018, Andrews_Webb_2018}.   
The influence of a pattern in the surface temperature is communicated to the TOA radiative fluxes by the clouds 
(Gregory and Andrews, 2016; Rose et al.; Silvers et al., 2018).  Thus, clouds are important factors to both the uncertainty 
estimates of climate sensitivity, as well as to the variability in time of the climate sensitivity.

A second motivation for this paper is to document the simulation of clouds in the newly developed GFDL models that 
have participated in CMIP6.  This will largely be done in the context of the \textit{amip} experiment and comparisons to observations
that overlap with parts of the textit{amip} time period (1979-2014).  The experiments and diagnostics used are taken from the latest 
round of CFMIP which has provided a useful framework with which to evaluate cloud feedbacks.

This work assesses the fidelity of the clouds in AM4.0/CM4.0 with observations and compares clouds from experiments 
with differing sensitivities. 
Section two briefly outlines the data used and the methods of analysis, with further details presented in the appendix.  Section
three compares the clouds from the \textit{amip} run of AM4.0 with satellite observations.  That section also 
analyzes the radiative fluxes at the top-of-the-atmosphere (TOA) and compares them to the observations from 
CERES-EBAF.    Section four documents the climate sensitivities of GFDL models across multiple CMIP6 experiments 
and discusses the potential role played by clouds.   Conclusions are drawn in section five.

@@

The comparisons with observations are made mostly with the \textit{amip} experiment that serves as one of the DECK CMIP6 
experiments (\cite{Eyring_etal_2016},  \cite{Webb_etal_2017}).   The experiments used 
in this paper are all part of the GFDL contribution to CMIP6 and as a result, the \textit{amip} and \textit{historical}
experiments end in the year 2014.  In addition to the \textit{amip} experiment we use the \textit{historical, abrupt-4xCO2}, and \textit{1pctCO2}
experiments from the CMIP6 DECK.  Several experiments from the CFMIP contributions 
to CMIP6 are also utilized.  These experiments include \textit{amip, amip-p4K, amip-m4K, amip-future4K, aqua-control} and \textit{aqua-p4K}.  
In order to utilize as much of the available data as possible much of the analysis of the \textit{amip} experiment uses the 
full span of years (1979-2014) despite the fact that the satellite observations do not extend as far back in time.  However, 
some of the analysis for the \textit{amip} and \textit{historical} experiments, in particular for the calculation of a climate feedback parameter over recent years, a shorter period (1996-2014) is used in order to eliminate complications due to radiative forcing in the early 1990's due to volcanic activity.  The relevant time period is stated in the captions of the figures and tables.

We compare output from the AM4.0 GCM \textit{amip} experiment to observations from ISCCP (the International Satellite Cloud Climatology Project), MISR (the Multiangle Imaging Spectro-Radiometer), MODIS (the Moderate Resolution Imaging Spectroradiometer), 
CALIPSO (the Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation; 
Chepfer et al., 2010; Cesana and Chepfer, 2013), and CERES-EBAF.
More details are provided in the appendix.  The CMIP6 and CFMIP experiments used in this study are given in Table
\ref{tab:exps} and the observational data sets used are given in Table \ref{tab:obs}.  Additional details about all 
experiments used in this paper are given in \cite{Eyring_etal_2016} and \cite{Webb_etal_2017}.  The experiments are referred to with the
appropriate CMIP6 labels.

Throughout the paper we utilize what has come to be known as the Cloud Radiative Effect (CRE), calculated 
as the difference between the full-sky and clear-sky radiative fluxes.  
The CRE is a simple and effective tool for evaluating the influence of 
clouds on the climate.  Note that CRE is often referred to as the Cloud Radiative Forcing (CRF) in earlier literature \cite<e.g.,>[]{Cess_etal_1989, Soden_etal_2004}.  It is well known (Soden et al., 2004) that a change in the CRE is not a direct measure of the cloud 
feedback, and that the CRE can be influenced by non-cloud factors.  However, there are many examples of 
the productive use of the CRE as a way to diagnose cloud effects (e.g. Cess et al., 1989; Webb et al., 2001; 
\citeA{Ringer_etal_2014}; Andrews et al., 2015), it is simple to calculate, and
it allows for direct comparisons with previous studies.  Cloud feedbacks calculated with either the so-called 
`partial radiative purturbation' or `radiative kernels' method correspond
to a positive shift (on the order of 0.5 $\rm{W m^{-2}K^{-1}}$) of the change in CRE values (Soden et al., 2004; Soden et al., 2008).  
The total cloud feedback of a model is therefore usually more positive than will be implied by the CRE.

@@

Before discussing the climate sensitivities of AM4/CM4 and how they are influenced by cloud feedbacks we assess
the clouds and radiative fluxes which are simulated by AM4.0 over the recent period of 1979-2014.  This is done 
by comparing satellite based observations to the \textit{amip} experiment that is part of GFDLs 
contribution to CMIP6.  
During the development and documentation of AM4.0 \cite{Zhao_etal_2018a, Zhao_etal_2018b}, 
(Z18a and Z18b hereafter) 
and CM4 \cite{Held_etal_2019}, specific cloud properties such as the 
cloud droplet number, cloud fraction, and the partition of cloud water into liquid or ice were noted, but 
accurately modeling the radiative fluxes at the top of the atmosphere was determined to be a 
higher priority partly because of the large observational uncertainties of those cloud properties (Z18a,b).   
Based on the initial analysis for those papers, the cloud properties of AM4.0 were determined to be
at least as consistent with observations as previous generations of GFDL models.  Z18a also noted 
that AM4.0 underestimates the liquid droplet number over oceans and downwind of the aerosol emission 
sources.   
Comparisons of AM4.0 with observations from ISCCP and MODIS in Z18a (their Figure 10) showed an underestimate of 
the optically thin low clouds and the mid-level clouds.  
However, the observational uncertainty for some of these properties prohibited firm conclusions.  This paper
uses additional observations and a more thorough analysis of the clouds in AM4/CM4 to better determine 
the quality of the cloud simulations.        
In this section we first compare the simulated and observed cloud fraction.  We then compare  
the simulated TOA radiative fluxes from AM4.0 to those observed by CERES.

@@

\subsection{COSP: change title}
In order to compare the observed and modeled clouds as consistently as possible across multiple platforms 
GFDL has implemented a collection of satellite simulators which are part of the 
CFMIP (Cloud Feedback Model Intercomparison Project; see \cite{Webb_etal_2017}).  Satellite simulator
are software tools that are used in coordination with GCMs to allow for model variables 
to be translated to variables that are more directly comparable with what satellites observe.  
These simulators are often referred to as the CFMIP Observation Simulator Package (COSP; see Bodas-Salcedo et al. 2011).  
This paper highlights simulators of the algorithms for 
four satellite datasets: ISCCP, 
MODIS, 
MISR, and 
CALIPSO.  By using multiple complimentary satellite 
datasets and simulators we can benefit from the different strengths of each satellite while largely 
minimizing the known problems and discrepancies of the observational data sets.

@@

The MISR satellite provides good observations of low-level cloud heights and amounts.  
For this reason and for comparison with both Kay et al., 2012, and Zhang et al., 2019 we compare the cloud fraction 
between observations from MISR and the AM4.0 simulator output for clouds below 3km in Figure \ref{fig:cf_misr_lowcl}.  
The optical depth is used to distinguish between the thick and thin clouds (optical depth greater than (thick), or less than (thin) 3.6).  
Consistent with comparisons with ISCCP and CALIPSO we see that AM4.0 has fewer clouds than observed.
For thick low-level clouds the negative cloud fraction bias is about 14\%, and 12\% for thin clouds.   
For thick clouds AM4.0 does reasonably well producing the observed spatial distribution of clouds, except in the tropical Indian Ocean. 
The observed pattern of thin clouds is not reproduced as well by the model in the tropics, and this is largely due to the 
presence of small cumulus.

@@

\textit{There have been several papers that suggest that changes in cloud phase have led to an increased shortwave cloud feedback in CMIP6 models (Flynn and Mauritsen, 2020).  With the exception of the aquaplanet experiment, all the experiments examined in this paper have identical microphysics parameterizations and tuning.  So the changes in high-latitude cloud feedback and distribution must be due to differences in boundary conditions or the atmospheric circulations.}
